Comparison of dimensionality reduction schemes for parallel global optimization algorithms \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: drop unused references

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}
%\setlength{\abovedisplayskip}{3pt}
%\setlength{\belowdisplayskip}{3pt}

\begin{document}
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization algorithms
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} }
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems. Algorithms are developed within the framework of the information-statistical approach and implemented in a parallel solver “Globalizer” . The optimization problem is solved by reducing the multidimensional problem to a set of joint one-dimensional problems that are solved in parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The results of computational experiments carried out on several hundred test problems are discussed.

@@

%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer\cite{globalizerSystem}, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
\par
Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is a time-consuming one. As a result, the overall time of solving the optimization
problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
is highly important, since an estimate of the global minimum can be constructed on the
basis of a finite number of computed values of the optimized function only in this case .

@@

\bibitem{censorZeniosParGO}% (MR1486040)
\newblock Y. Censor and S. A. Zenios,
\newblock \emph{Parallel Optimization: Theory, Algorithms, and Applications},
\newblock Oxford University Press, 1998.

\bibitem{ciegisHentyParGO} (MR2499546)% [10.1007/978-0-387-09707-7]
\newblock R. \v Ciegis, D. Henty, B. K\aa gstr\"om and J. \v Zilinskas,
\newblock \emph{Parallel Scientific Computing and Optimization: Advances and Applications},
\newblock Springer, 2009.

\bibitem{Famularo1999} %(MR1827451) [10.1016/S0005-1098(99)00058-8]
\newblock D. Famularo, P. Pugliese and Y. D. Sergeyev,
\newblock \emph{\emph{A global optimization technique for checking parametric robustness}},
\newblock \emph{Automatica}, \textbf{35} (1999), 1605--1611.

\bibitem{fasanoPinter2013}% (MR3013800) [10.1007/978-1-4614-4469-5]
\newblock G. Fasano and J. D. Pint\'er,
\newblock \emph{Modeling and Optimization in Space Engineering},
\newblock Springer, 2013.

\bibitem{floudasPardalosGOState} %(MR1390521) [10.1007/978-1-4613-3437-8]
\newblock C. A. Floudas and M. P. Pardalos,
\newblock \emph{State of the Art in Global Optimization: Computational Methods and Applications},
\newblock Kluwer Academic Publishers, Dordrecht, 1996.

\bibitem{floudasPardGO}% (MR1147432) [10.1007/s10898-008-9332-8]
\newblock C. A. Floudas and M. P. Pardalos,
\newblock \emph{Recent Advances in Global Optimization},
\newblock Princeton University Press, 2016.

\bibitem{Gablonsky} %(MR1856800) [10.1023/A:1017930332101]
\newblock J. M. Gablonsky and C. T. Kelley,
\newblock \emph{\emph{A locally-biased form of the DIRECT algorithm}},
\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 27--37.

\bibitem{horstTuyGO}% (MR1102239) [10.1007/978-3-662-03199-5]
\newblock R. Horst and H. Tuy,
\newblock \emph{Global Optimization: Deterministic Approaches},
\newblock Springer-Verlag, Berlin, 1990.

\bibitem{Kvasov2015}% (MR3585540) [10.1016/j.advengsoft.2014.09.014]
\newblock D. E. Kvasov and Y. D. Sergeyev,
\newblock \emph{\emph{Deterministic approaches for solving practical black-box global optimization problems}},
\newblock \emph{Adv. Eng. Softw.}, \textbf{80} (2015), 58--66.

\bibitem{Menniti}% [10.1016/j.epsr.2007.10.009]
\newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Y. D. Sergeyev and N. Sorrentino,
\newblock \emph{\emph{Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on
 efficient domain partitions}},
\newblock \emph{Electric Power Systems Research}, \textbf{78} (2008), 1217--1229.

\bibitem{locatelliSchoenGO}% (MR3136805) [10.1137/1.9781611972672]
\newblock M. Locatelli and F. Schoen,
\newblock \emph{Global Optimization: Theory, Algorithms and Applications},
\newblock SIAM, 2013.

\bibitem{luqueAlbaGA}% (MR2808873) [10.1007/978-3-642-22084-5]
\newblock G. Luque and E. Alba,
\newblock \emph{Parallel Genetic Algorithms. Theory and Real World Applications},
\newblock Springer-Verlag, Berlin, 2011.

\bibitem{pardalosZhigljavskyZilinskas2016}% (MR2361744) [10.1007/978-3-319-29975-4]
\newblock M. P. Pardalos, A. A. Zhigljavsky and J. \v Zilinskas,
\newblock \emph{Advances in Stochastic and Deterministic Global Optimization},
\newblock Springer, 2016.

\bibitem{pinterGO}% (MR1374104) [10.1007/978-1-4757-2502-5]
\newblock J. D. Pint\'er,
\newblock \emph{Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications)},
\newblock Kluwer Academic Publishers, Dordrecht, 1996.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}

\begin{document}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization algorithms
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} }
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems. Algorithms are developed within the framework of the information-statistical approach and implemented in a parallel solver “Globalizer” . The optimization problem is solved by reducing the multidimensional problem to a set of joint one-dimensional problems that are solved in parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The results of computational experiments carried out on several hundred test problems are discussed.

@@

%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer\cite{globalizerSystem}, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\) space.
\par
Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is a time-consuming one. As a result, the overall time of solving the optimization
problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
is highly important, since an estimate of the global minimum can be constructed on the
basis of a finite number of computed values of the optimized function only in this case .

@@
